---
title: "Spatial Modeling HW"
author: "Daniela Adjunta"
date: "2/12/2019"
output: html_document
---

```{r}
library(vegan)
data("BCI")
BCI_xy = data.frame(x = rep(seq(625754, 626654, by=100), each=5), 
                    y = rep(seq(1011569,  1011969, by=100), len=50))
```
1.
```{r}
colSums(BCI)
```
For the analysis of a rare and common species I used Virola sebifera (617 observed) as my common species and Sorocea affinis (28 observed) as my rare species. 
```{r}
vir_dist = dist(BCI$Virola.sebifera)
sor_dist = dist(BCI$Sorocea.affinis)
xy_dist = dist(BCI_xy)
max_dist = max(xy_dist) / 2
```
Virola sebifera - common species 
```{r}
plot(xy_dist , vir_dist)
abline(lm(vir_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, vir_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)
obs_corv = cor(xy_dist, vir_dist)
obs_corv
vir_mantel = mantel(xy_dist, vir_dist)
vir_mantel
```
Sorocea affinis - rare species 
```{r}
plot(xy_dist , sor_dist)
abline(lm(sor_dist ~ xy_dist), lwd=3, col='red')
lines(lowess(xy_dist, sor_dist), lwd=3, col='pink')
abline(v = max_dist, col='red', lwd=3, lty=2)
obs_cors = cor(xy_dist, sor_dist)
obs_cors
sor_mantel = mantel(xy_dist, sor_dist)
sor_mantel
```

2. 
```{r}
sp_ids = c("Drypetes.standleyi", "Cordia.lasiocalyx", "Hirtella.triandra",
           "Picramnia.latifolia", "Quassia.amara",
           "Tabernaemontana.arborea", "Trattinnickia.aspera", 
           "Xylopia.macrantha")
```
above I added Drypetes.standleyi to the subset, so I could have a subset that I can compare this species to all the others in the subset
```{r}
x <- BCI_xy$x
y <- BCI_xy$y
d_1var_lm = gls(Drypetes.standleyi ~ Cordia.lasiocalyx, data = BCI)
par(mfrow=c(1,1))
plot(Variogram(d_1var_lm, form= ~ x + y))
d_1var_er = update(d_1var_lm, corr=corExp(form = ~ x + y))
plot(Variogram(d_1var_er, form= ~ x + y))
anova(d_1var_lm, d_1var_er)
```
```{r}
abun_dat = data.frame(BCI, BCI_xy)
BCI_sp = subset(abun_dat, select=sp_ids)
d_mult_lm = gls(Drypetes.standleyi ~ ., data = BCI_sp)
plot(Variogram(d_1var_lm, form= ~ x + y))
d_mult_er = update(d_mult_lm, corr=corExp(form = ~ x + y))
plot(Variogram(d_mult_er, form= ~ x + y))
anova(d_mult_lm, d_mult_er)
```
What was very interesting about this analysis was that the plots before using the error structures seem to be identical for both the comparison of one species to multiple. However after the use of the error structure, differences were clear. Therefore, the use of the error structure was very important in terms of understanding the data. It defineitly improved the model fit, you can really tell by looking at the plotted points in comparion to the trendline. 